### 2024-05-31 2024-5-13/sections/heuristic.tex
section 7.1

Let us describe the optimization more precisely. 
\begin{itemize}
\item For a regular expression $x \in (e_1)^{\{m,n\}}$ where $e_1$ contains counting operators (but does not contain the nesting of counting operators), we can unfold the counting operators in the following ways to obtain a register-representable RECL constraint: (i) unfold the counting operators in $e_1$, (ii) unfold the counting operator $^{\{m,n\}}$. 
%
Let $e', e''$ denote the resulting constraints.

\item We estimate the sizes of the two CEFAs $\aut', \aut''$ constructed from $e', e''$ respectively as pairs $(s', r')$ and $(s'', r'')$, where $s'$ and $r'$ are the estimates of the number of states and registers in $\aut'$, similarly for $s''$ and $r''$. 
%
\item From the construction in Section~\ref{sec:algorithm}, 
two NFAs $\cU_{\aut'}, \cU_{\aut''}$ are constructed from $\aut', \aut''$ by dropping the accepting conditions and ignoring the characters. As a result, the alphabets of $\cU_{\aut'}, \cU_{\aut''}$ are the set of integer vectors  $\myvec{v}$ occurring in the transitions of $\aut', \aut''$ respectively. The integer vectors of  $\aut'$ are of form $(\underbrace{0, \cdots, 0}_{i-1}, 1, \underbrace{0, \cdots, 0}_{r'-i-1})$, where $i \in [r']$. Consequently, the size of the alphabet of $\cU_{\aut'}$ is bounded by $r'$. Similarly for $\cU_{\aut''}$. 
Finally, we estimate the sizes (number of transitions) of the NFAs $\cU_{\aut'}, \cU_{\aut''}$ as $sz' = (s')^2 r'$ and $sz''= (s'')^2 r''$ respectively.
%
\item If $sz' \le sz''$, then we choose to unfold the counting operators in $e_1$. Otherwise, we choose to unfold the counting operator $^{\{m,n\}}$. 
\end{itemize}

In the sequel, we show how we estimate the sizes of CEFAs constructed from register-representable regular expressions.

Let $\regex$ be a register-representable regular expression. Then we compute $(s_\regex, r_\regex) \in \Nat \times \Nat$, i.e. the pair (estimate of the number of states, estimate of the number of registers), as follows. 
\begin{itemize}
  \item If $\regex = \emptyset$ or $\regex = \epsilon$, then $s_{\regex} = 0$ and $r_\regex = 0$. 
  \item If $\regex = \regex_1 \concat \regex_2$, then $s_{\regex} = s_{\regex_1} + s_{\regex_2}$ and $r_{\regex} = r_{\regex_1} + r_{\regex_2}$.
  \item If $\regex = \regex_1 + \regex_2$, then $s_{\regex} = s_{\regex_1} + s_{\regex_2}$ and $r_{\regex} = r_{\regex_1} + r_{\regex_2}+1$. 
  \item If $\regex = \regex_1 \cap \regex_2$, then $size(\regex) = size(\regex_1)*size(\regex_2)$ and $r_{\regex} = r_{\regex_1} + r_{\regex_2}$.
  \item If $\regex = \overline{\regex_1}$, then $s_\regex = 2^{s_{\regex_1}}$ (since the determinization is needed for the complement) and $r_\regex = 0$.
  \item If $\regex = \regex_1^*$, then $s_{\regex} = s_{\regex_1}$ and $r_{\regex} = 0$.
%  \item $\regex = \regex_1 \backslash \regex_2$ can be transferred to $\regex = \regex_1 \cap \overline{\regex_2}$ and then the size can be computed by the formula of $\cap$ and $\bar{}$.
  \item For $\regex = \regex_1^{\{m, n\}}$ or $\regex = \regex_1^{\{n,\infty\}}$, $s_{\regex} = s_{\regex_1}$ and $r_{\regex} = 1$.
\end{itemize}

@@

\medskip
\noindent
\emph{Step 1. (Compute the number of states generated by each counting operator if we unwind it).}

@@

\begin{example}
  Consider the regex $(a^{\{1,1000\}})^{\{1,2\}}$. The score of $^{\{1,1000\}}$ is $1000*(0+1) = 1000$ and the score of $^{\{1,2\}}$ is $2*(2+1) = 6$. We unwind the outer counting operator first and obtain a CEFA containing only 2 states.  
\end{example}

>>>>

Let us describe the optimization more precisely. 
\begin{itemize}
\item For a regular expression $e = (e_1)^{\{m,n\}}$ where $e_1$ is a register-representable regular expression, we can unfold the counting operators in the following ways to obtain a register-representable RECL constraint: (i) unfold the counting operators in $e_1$, (ii) unfold the counting operator $^{\{m,n\}}$. 
%
Let $\ufld_1(e)$ and $\ufld_2(e)$ denote the resulting regular expressions.

\item We estimate the sizes of the two CEFAs $\aut_{\ufld_1(e)}, \aut_{\ufld_2(e)}$ constructed from $\ufld_1(e), \ufld_2(e)$ respectively as pairs $(\snum_1(e), \rnum_1(e))$ and $(\snum_2(e), \rnum_2(e))$, where $\snum_1(e)$ and $\rnum_1(e)$ are the estimates of the number of states and registers in $\aut_{\ufld_1(e)}$, similarly for $(\snum_2(e)$ and $\rnum_2(e))$. 
%
\item From the construction in Section~\ref{sec:algorithm}, 
two NFAs $\cU_{\aut_{\ufld_2(e)}}, \cU_{\aut_{\ufld_2(e)}}$ are constructed from $\aut_{\ufld_1(e)}$ and $\aut_{\ufld_2(e)}$ by dropping the accepting conditions and ignoring the characters. As a result, the alphabets of $\cU_{\aut_{\ufld_2(e)}}, \cU_{\aut_{\ufld_2(e)}}$ are the set of integer vectors  $\myvec{v}$ occurring in the transitions of $\aut_{\ufld_1(e)}, \aut_{\ufld_2(e)}$ respectively. The integer vectors of  $\aut_{\ufld_1(e)}$ are of form $(\underbrace{0, \cdots, 0}_{i-1}, 1, \underbrace{0, \cdots, 0}_{r'-i-1})$, where $i \in [r']$. Consequently, the size of the alphabet of $\cU_{\aut_{\ufld_1(e)}}$ is bounded by $\rnum_1(e)$. Similarly for $\cU_{\aut_{\ufld_1(e)}}$. 
Finally, we estimate the sizes (number of transitions) of the NFAs $\cU_{\aut_{\ufld_1(e)}}, \cU_{\aut_{\ufld_2(e)}}$ as $sz_1 = (\snum_1(e))^2 \cdot \rnum_1(e)$ and $sz_2 = (\snum_2(e))^2 \cdot \rnum_2(e)$ respectively.
%
\item If $sz_1 \le sz_2$, then we choose to unfold the counting operators in $e_1$. Otherwise, we choose to unfold the counting operator $^{\{m,n\}}$. 
\end{itemize}

%In the sequel, we show how we estimate the sizes of CEFAs constructed from register-representable regular expressions.

It remains to show how to compute the pairs $(\snum_1(e), \rnum_1(e))$ and $(\snum_2(e), \rnum_2(e))$ from $e=(e_1)^{\{m,n\}}$. Let $e=(e_1)^{\{m,n\}}$ be a regular expression where $e_1$ is a register-representable regular expression. Then we compute $\snum'_1(e_1)$ and $(\snum'_2(e_1), \rnum'_2(e_1))$ by a structural induction on $e_1$. Finally, let $(\snum_1(e), \rnum_1(e)) = (\snum'_1(e_1), 1)$ and $(\snum_2(e), \rnum_2(e)) = (\snum'_2(e_1), \rnum'_2(e_1))$.
\begin{itemize}
  \item If $e_1 = \emptyset$ or $e_1 = \epsilon$, then $\snum'_1(e_1)  = 0$ and $(\snum'_2(e_1), \rnum'_2(e_1))  = (0, 0)$. 
%
 %
  \item If $e_1 = e_2 \concat e_3$, then $\snum'_1(e_1) = \snum'_1(e_2) + \snum'_1(e_3)$ and 
%
  $$
  \begin{array}{l }
  (\snum'_2(e_1), \rnum'_2(e_1)) =  \\
  \ \ \ \ (n(\snum'_2(e_2) + \snum'_2(e_3)), n(\rnum'_2(e_2) + \rnum'_2(e_3))).
 \end{array}
  $$
%
 \item If $e_1 = e_2 + e_3$, then $\snum'_1(e_1) = \snum'_1(e_2) + \snum'_1(e_3)$ and 
%
  $$
  \begin{array}{l }
  (\snum'_2(e_1), \rnum'_2(e_1)) =  \\
  \ \ \ \ (n(\snum'_2(e_2) + \snum'_2(e_3)), n(\rnum'_2(e_2) + \rnum'_2(e_3)+1)).
 \end{array}
  $$
  (Recall that the construction of $\aut_1 \cup \aut_2$ introduces one additional register.)
%
 \item If $e_1 = e_2 \cap e_3$, then $\snum'_1(e_1) = \snum'_1(e_2) \times \snum'_1(e_3)$ and 
%
  $$
  \begin{array}{l }
  (\snum'_2(e_1), \rnum'_2(e_1)) =  \\
  \ \ \ \ (n(\snum'_2(e_2) \times \snum'_2(e_3)), n(\rnum'_2(e_2) + \rnum'_2(e_3))).
 \end{array}
  $$
%
 \item If $e_1 = \overline{e_2}$, then $\snum'_1(e_1) = 2^{\snum'_1(e_2)}$ and 
%
  $
%  \begin{array}{l }
  (\snum'_2(e_1), \rnum'_2(e_1)) =  (n \cdot 2^{\snum'_2(e_2)}, 0).
% \end{array}
  $
 (The exponential blow-up is explained by the fact that the determinization is needed for the complementation.) 
%
%  \item If $\regex = \overline{\regex_1}$, then $s_\regex = 2^{s_{\regex_1}}$ (since the determinization is needed for the complement) and $r_\regex = 0$.
%
 \item If $e_1 = (e_2)^*$, then $\snum'_1(e_1) = \snum'_1(e_2)$ and 
%
  $
%  \begin{array}{l }
  (\snum'_2(e_1), \rnum'_2(e_1)) =  (n \snum'_2(e_2), 0).
% \end{array}
  $
%  \item If $\regex = \regex_1^*$, then $s_{\regex} = s_{\regex_1}$ and $r_{\regex} = 0$.
%  \item $\regex = \regex_1 \backslash \regex_2$ can be transferred to $\regex = \regex_1 \cap \overline{\regex_2}$ and then the size can be computed by the formula of $\cap$ and $\bar{}$.
 \item If $e_1 = (e_2)^{\{m', n'\}}$ or $e_1 = (e_2)^{\{n', \infty\}}$, then $\snum'_1(e_1) = n' \snum'_1(e_2)$ and 
%
  $
%  \begin{array}{l }
  (\snum'_2(e_1), \rnum'_2(e_1)) =  (n \snum'_2(e_2), n).
% \end{array}
  $
%  \item For $\regex = \regex_1^{\{m, n\}}$ or $\regex = \regex_1^{\{n,\infty\}}$, $s_{\regex} = s_{\regex_1}$ and $r_{\regex} = 1$.
\end{itemize}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\hide{
\medskip
\noindent
\emph{Step 1. (Compute the number of states generated by each counting operator if we unwind it).}

@@

\begin{example}
  Consider the regex $(a^{\{1,1000\}})^{\{1,2\}}$. The score of $^{\{1,1000\}}$ is $1000*(0+1) = 1000$ and the score of $^{\{1,2\}}$ is $2*(2+1) = 6$. We unwind the outer counting operator first and obtain a CEFA containing only 2 states.  
\end{example}
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
